---
title: "dplyr & purrr を用いたデータハンドリング"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
 pdf_document:
  latex_engine: xelatex
  fig_width: 5
  fig_height: 4
  number_sections: true
  toc: true
  toc_depth: 2
graphics: yes
mainfont: YuMincho
monofont: Ricty Discord
---
\fontsize{9}{12}
\hrulefill

This exercise is taken from [dplyr & purrr を用いたデータハンドリング](https://www.slideshare.net/SomatoriKeita/dplyr-purrr-77749974). Original data is on [e-Stat: 精神科病院の推計患者数，年齢階級 × 性・疾病分類（精神及び行動の障害） × 入院－外来別](https://www.e-stat.go.jp/SG1/estat/GL08020103.do?_toGL08020103_&listID=000001141596&requestSender=dsearch)

# Load Libraries
```{r, echo=F, warning=FALSE, message=FALSE}
setwd("~/temp/")
```
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
```

# Read Data
Try reading data without any option. We can see untidy data.
```{r, warning=FALSE, message=FALSE}
data <- readr::read_csv("j0056.csv",
                        locale = readr::locale(encoding="cp932")
                        )
knitr::kable(data[1:5, 1:3])
```

Let's skip first two lines to make it better.
```{r, warning=FALSE, message=FALSE}
data <- readr::read_csv("j0056.csv",
                        locale = readr::locale(encoding="cp932"),
                        skip=2)
colnames(data)
```

## Make tidy data

```{r, warning=FALSE, message=FALSE}
data %>% 
    fill(X1) %>% # fill X1 with previous value, here "男"
    slice(46:63) %>%
    filter(X2 != "総数") %>%  # remove row with "総数" in column "X2"
    filter(X2 != "（再掲）精神及び⾏動の障害") %>%
    select(-3) %>% # delete the third column
    select(-23:-26) %>%
    rename(sex=X1, disease_name=X2) -> data
knitr::kable(data[1:5, 1:5])
```

Make it tidier:

```{r, warning=FALSE, message=FALSE}
data %>% gather(`　　　０歳`:`９０歳以上`, # specify columns
                key=age, # original column names will be in a new column named "age"
                value="count" # original values will be in "count"
               ) -> data 
knitr::kable(data[1:10, 1:4])
```

Final modification:

```{r, warning=FALSE, message=FALSE}
data %>%
  rowwise() %>%
  mutate(count = ifelse(count=="-", NA, count), # create NA
         disease_name = stringr::str_trim(disease_name, side="both") # trim spaces
        ) %>%
  mutate(count = as.numeric(count)) -> data
knitr::kable(data[1:10, 1:4])
```

# Transform data

```{r, warning=FALSE, message=FALSE}
schizo <- data %>% 
            filter(disease_name == "統合失調症，統合失調症型障害及び妄想性障害",
                   !is.na(count)) # do not take rows with NA
```

# Visualize

```{r, warning=FALSE, message=FALSE, dev="quartz_pdf"}
ggplot(data=schizo) +
  geom_bar(aes(x=age, y=count), stat="identity") +
  scale_y_continuous(expand=c(0,0)) +
  theme_gray(base_family = "YuGo-Medium")
```

# purrr

```{r, warning=FALSE, message=FALSE}
mapped <- data %>%
  split(list(.$disease_name, .$sex)) %>% # make list by name and sex
  map(group_by, age) %>% # applying a function to the all lists
  map(summarize, mean=mean(count, na.rm=T))
```

You can apply self-defined functions:

```{r, warning=FALSE, message=FALSE}
myhist <- function(data){
  ggplot(data) +
  geom_bar(aes(x=age, y=mean), stat="identity") +
  scale_y_continuous(expand=c(0,0)) +
  theme_gray(base_family = "YuGo-Medium")
}

mappedhist <- data %>%
  split(list(.$disease_name, .$sex)) %>% # make list by name and sex
  map(group_by, age) %>% # applying a function to the all lists
  map(summarize, mean=mean(count, na.rm=T)) %>%
  map(myhist)
names(mappedhist)
```


You can take out any figure:

```{r, warning=FALSE, dev="quartz_pdf", message=FALSE}
mappedhist$血管性及び詳細不明の認知症.男
```

You can make a list of analysis:

```{r, eval=FALSE, warning=FALSE, message=FALSE}
data %>% split(.$disease_name) %>%
  map(~lm(count ~ sex, data=.)) %>%
  map(summary) %>%
  map("coefficients")
```

# Map family

* Change return values
 + `map_lg()`
 + `map_chr()`
 + `map_int()`
 + `map_dbl()`
* map multiple lists
 + `map2()`
 + `map3()`
 + `map_n()`
